---
title: "10x vs Plate"
output:
  html_document:
    df_print: paged
---

```{r}
#install.packages("tidyverse")
library(tidyverse)
library(stringr)
library(Seurat)

Sys.setenv(R_MAX_NUM_DLLS=150)
```

```{r}
load("~/maca/save/All_seurat_tiss.Robj")
load("~/maca/save/10x_All_seurat_tiss.Robj")

tiss@meta.data['cell'] <- rownames(tiss@meta.data)
tissX@meta.data['cell'] <- rownames(tissX@meta.data)

```

Add in percent Rn54s.
```{r}
percent.Rn45s <- Matrix::colSums(tiss@raw.data[c('Rn45s'), ])/Matrix::colSums(tiss@raw.data)
tiss <- AddMetaData(object = tiss, metadata = percent.Rn45s, col.name = "percent.Rn45s")

percent.Rn45s <- tissX@raw.data[c('Rn45s'), ]/Matrix::colSums(tissX@raw.data)
tissX <- AddMetaData(object = tissX, metadata = percent.Rn45s, col.name = "percent.Rn45s")
```

Extract just the metadata from plates and tenx cells.
```{r}
just_plates = as_tibble(tiss@meta.data[c('tissue', 'subtissue', 'percent.ribo', 'percent.Rn45s','nGene', 'plate.barcode', 'nReads')])
just_plates <- rownames_to_column(just_plates, "cell")
just_plates['method'] <- 'plate'
just_plates <- rename(just_plates, run = plate.barcode)
just_plates <- rename(just_plates, nMolecules = nReads)

just_tenx = as_tibble(tissX@meta.data[c('tissue', 'percent.ribo', 'nGene','channel', 'percent.Rn45s', 'nUMI')])
just_tenx <- rownames_to_column(just_tenx, "cell")
just_tenx <- rename(just_tenx, nMolecules = nUMI)

just_tenx['method'] <- '10x'
just_tenx <- rename(just_tenx, run = channel)

df <- bind_rows(just_plates, just_tenx)
```

Rename mammary from tenx and check tissue counts.

```{r}
df <- df %>% mutate(tissue = fct_recode(tissue,
    "Mammary_Gland" = "Mammary"))
df %>% count(tissue)
```

```{r}
common_tissues <- as.character(df %>% 
                                 group_by(tissue, method) %>% summarize(count = n()) %>% 
                                 ungroup() %>% group_by(tissue) %>% summarize(count = n()) %>% 
                                 filter(count > 1) %>% pull(tissue))
common_tissues
```

## Differences in genes and ribo fraction for plates and 10x.

```{r, fig.width = 12, fig.height = 8}
filter(df, tissue %in% common_tissues) %>% ggplot(aes(nGene, ..density..,colour = method)) + 
  geom_freqpoly(binwidth = 300) +
  facet_wrap(~ tissue, ncol = 4) + 
  ggtitle("Number of genes expressed")
```



```{r, fig.width = 12, fig.height = 8}
filter(df, tissue %in% common_tissues) %>% ggplot(aes(percent.ribo, ..density..,colour = method)) + 
  geom_freqpoly(binwidth = 0.02) +
  facet_wrap(~ tissue, ncol = 4) + 
  ggtitle("Percent ribosomal protein")
```

```{r, fig.width = 12, fig.height = 8}
filter(df, tissue %in% common_tissues) %>% ggplot(aes(percent.Rn45s, ..density..,colour = method)) + 
  geom_freqpoly(binwidth = 0.02) +
  facet_wrap(~ tissue, ncol = 4) + 
  ggtitle("Percent Rn45s")
```

The enrichment of Rn45s and ribosomal genes differ strongly between the two methods.

```{r}
tissue_colors <- read_csv('~/maca/metadata/tissue_colors.csv')
tissue_colors <- rename(tissue_colors, tissue = X1)
```

```{r}
medians <- df %>% group_by(method, tissue, run) %>% 
                  summarize(median_Rn45s = median(percent.Rn45s), 
                            median_ribo = median(percent.ribo),
                            n_cells = n()) 
medians <- left_join(medians, tissue_colors, by = 'tissue')

medians %>% filter(tissue %in% common_tissues,) %>% ggplot(aes(x = median_Rn45s, y = median_ribo, color = color, shape = method)) + geom_point()  +
  scale_color_identity(breaks = tissue_colors$color,
                       labels = tissue_colors$tissue,
                       guide = "legend") + 
  ggtitle("Comparing Rn45s and ribo genes between channels and plates.")
```




##  Dynamic Range of individual cells

We compare individual cells from tissues containing similar numbers of genes per cell.


Get a few cells from each method with values in a certain range.
```{r}
subset.size <- 1
set.seed(1)
cell.subset <- filter(df, tissue == 'Heart', nGene > 2900, nGene < 3100) %>% 
    group_by(method) %>% 
    sample_n(subset.size) %>%
    pull(cell)
```

It's difficult to compare UMI and reads apples to apples. Here both are scaled to 10k reads/UMIs then log(1+x)'d.

```{r, fig.width = 4, fig.height = 8}
par(mfrow=c(2,1))

expr <- tissX@data[,cell.subset[1]]
expr <- expr[expr > 0]
plot(sort(expr), ylab = "log expression", xlab = "gene index", col = "red")
title(main = "Heart 10x cell")

expr <- tiss@data[,cell.subset[2]]
expr <- expr[expr > 0]
plot(sort(expr), ylab = "log expression", xlab = "gene index", col = "green")
title(main = "Heart plateseq cell")
```

If we instead scaled them to 10^6 both, we would see a different shape for the plate data, with more dynamic range in the low values.

```{r, fig.width = 4, fig.height = 8}
par(mfrow=c(2,1))

expr <- adjust_lognorm(tissX@data[,cell.subset[1]], 100)
expr <- expr[expr > 0]
plot(sort(expr), ylab = "log expression", xlab = "gene index", col = "red")
title(main = "Heart 10x cell")

expr <- adjust_lognorm(tiss@data[,cell.subset[2]], 100)
expr <- expr[expr > 0]
plot(sort(expr), ylab = "log expression", xlab = "gene index", col = "green")
title(main = "Heart plateseq cell")
```


## Comparing the Average T cell


```{r}
anno_plate = read_csv('~/maca/metadata/maca_3month_combined_cell_annotations.csv')
anno_plate <- rename(anno_plate, cell = X1)
#anno_plate <- rename(anno, anno_tissue = tissue)
anno_plate %>% group_by(tissue, annotation, subannotation) %>% summarize(count = n())

anno_10x = read_csv('~/maca/metadata/maca_3month_combined_cell_annotations_10x.csv')
anno_10x <- rename(anno_10x, cell = X1)
#anno_plate <- rename(anno, anno_tissue = tissue)
anno_10x %>% group_by(tissue, annotation, subannotation) %>% summarize(count = n())
```

```{r}

cells_plate <- anno_plate %>% filter(annotation == 't_cells' & tissue == 'Spleen') %>% pull(cell)
pct_plate <- Matrix::rowMeans(tiss@raw.data[,cells_plate] > 0)

cells_10x <- anno_10x %>% filter(annotation == 't_cells' & tissue == 'Spleen') %>% pull(cell)
pct_10x <- Matrix::rowMeans(tissX@raw.data[,cells_10x] > 0)
```

```{r}
pct1 = as_data_frame(pct_10x)
pct1['gene'] <- names(pct_10x)
pct1['method'] <- 'tenx'
pct1 <- filter(pct1, value > 0.01)

pct2 = as_data_frame(pct_plate)
pct2['gene'] <- names(pct_plate)
pct2['method'] <- 'plate'
pct2 <- filter(pct2, value > 0.01)

pct <- bind_rows(pct1, pct2)
```



```{r}
pct %>% spread(key = method, value = value, fill = 0) %>% 
  ggplot(aes(x = plate, y = tenx)) + geom_point(alpha = 0.3) + geom_smooth() + geom_abline(slope=1, intercept=0) 
```


```{r}
library(scales)
```

```{r}
pct %>% spread(key = method, value = value, fill = 0) %>% 
  ggplot(aes(x = plate, y = tenx)) + geom_hex() + geom_abline(slope=1, intercept=0)  + scale_fill_continuous(trans = log10_trans())
```



```{r}
plate_v_tenx <- function(tissue_var, annotation_var){

 
cells_plate <- anno_plate %>% filter(annotation == annotation_var & tissue == tissue_var) %>% pull(cell)
pct_plate <- Matrix::rowMeans(tiss@raw.data[,cells_plate] > 0)
pct_plate
cells_10x <- anno_10x %>% filter(annotation == annotation_var & tissue == tissue_var) %>% pull(cell)
pct_10x <- Matrix::rowMeans(tissX@raw.data[,cells_10x] > 0)

pct1 = as_data_frame(pct_10x)
pct1['gene'] <- names(pct_10x)
pct1['method'] <- 'tenx'
pct1 <- filter(pct1, value > 0.01)

pct2 = as_data_frame(pct_plate)
pct2['gene'] <- names(pct_plate)
pct2['method'] <- 'plate'
pct2 <- filter(pct2, value > 0.01)

pct <- bind_rows(pct1, pct2)

pct  %>% spread(key = method, value = value, fill = 0) %>% 
  ggplot(aes(x = plate, y = tenx)) + geom_hex() + geom_abline(slope=1, intercept=0)  + scale_fill_continuous(trans = log10_trans())
}
```

## Individual cell types in tissues

```{r}
plate_v_tenx("Kidney", "tubule_cells")
```

```{r}
plate_v_tenx("Bladder", "luminal_cells")
```

```{r}
plate_v_tenx("Bladder", "basal_cells")
```

```{r}
plate_v_tenx("Liver", "hepatocytes")
```

```{r}
plate_v_tenx("Spleen", "t_cells")
```

```{r}
plate_v_tenx("Pancreas", "b")
```

```{r}
plate_v_tenx("Thymus", "t_cells")
```

### Global stats


Number 0 in plates and nonzero in 10x, and vice versa.
Fraction on left or right of the equality line. Average absolute percent difference. Conditioned on present in one at > 5%.


Side investigation:

Genex expressed in every cell of one of the methods but not in the other. This could be caused by spillover in droplets.
Best eg is t-cells in the thymus. Feels like qc.



## Miscellaneous




```{r}
df %>% group_by(annotation.y, tissue, plate.barcode.x) %>% summarize(count = n()) %>% ungroup() %>% group_by(annotation.y, tissue) %>% summarize(n_plates = n()) %>% arrange(desc(n_plates))

df %>% group_by(annotation.y, tissue) %>% summarize(count = n()) %>% ungroup() %>% group_by(annotation.y) %>% summarize(n_tissues = n()) %>% arrange(desc(n_tissues))
```



```{r}
df %>% group_by(annotation.y, subannotation.y) %>% summarize(count = n()) %>% ungroup()
```


```{r}
df %>% group_by(annotation.y, tissue, plate.barcode.x) %>% summarize(count = n()) %>% ungroup() %>% group_by(annotation.y, tissue) %>% summarize(n_plates = n()) %>% arrange(desc(n_plates))
```

