---
title: "tenx vs Plate"
output:
  html_document:
    df_print: paged
---

```{r}
library(tidyverse)
library(stringr)
library(Seurat)

Sys.setenv(R_MAX_NUM_DLLS=150)
```

```{r}
load("~/maca/save/All_seurat_tiss.Robj")

load("~/maca/save/10x_All_regressed_seurat_tiss.Robj")

tiss@meta.data['cell'] <- rownames(tiss@meta.data)
tissX@meta.data['cell'] <- rownames(tissX@meta.data)
```

Add in percent Rn54s.
```{r}
percent.Rn45s <- Matrix::colSums(tiss@raw.data[c('Rn45s'), ])/Matrix::colSums(tiss@raw.data)
tiss <- AddMetaData(object = tiss, metadata = percent.Rn45s, col.name = "percent.Rn45s")

percent.Rn45s <- tissX@raw.data[c('Rn45s'), ]/Matrix::colSums(tissX@raw.data)
tissX <- AddMetaData(object = tissX, metadata = percent.Rn45s, col.name = "percent.Rn45s")
```

Extract just the metadata from plates and tenx cells.
```{r}
just_plates = as_tibble(tiss@meta.data[c('tissue', 'subtissue', 'percent.ribo', 'percent.Rn45s','nGene', 'plate.barcode', 'nReads')])
just_plates <- rownames_to_column(just_plates, "cell")
just_plates['method'] <- 'plate'
just_plates <- rename(just_plates, run = plate.barcode)
just_plates <- rename(just_plates, nMolecules = nReads)

just_tenx = as_tibble(tissX@meta.data[c('tissue', 'percent.ribo', 'nGene','channel', 'percent.Rn45s', 'nUMI')])
just_tenx <- rownames_to_column(just_tenx, "cell")
just_tenx <- rename(just_tenx, nMolecules = nUMI)

just_tenx['method'] <- 'tenx'
just_tenx <- rename(just_tenx, run = channel)

df <- bind_rows(just_plates, just_tenx)
```

Rename mammary from tenx and check tissue counts.

```{r}
df <- df %>% mutate(tissue = fct_recode(tissue,
    "Mammary_Gland" = "Mammary"))
df %>% count(tissue)
```

```{r}
common_tissues <- as.character(df %>% 
                                 group_by(tissue, method) %>% summarize(count = n()) %>% 
                                 ungroup() %>% group_by(tissue) %>% summarize(count = n()) %>% 
                                 filter(count > 1) %>% pull(tissue))
common_tissues
```

## Differences in genes and ribo fraction for plates and tenx.

```{r, fig.width = 12, fig.height = 8}

filter(df, tissue %in% common_tissues) %>% ggplot(aes(nGene, ..density..,colour = method)) + 
  geom_freqpoly(binwidth = 300) +
  facet_wrap(~ tissue, ncol = 4) + 
  ggtitle("Number of genes expressed")

ggsave('figures/ngenes_comparison.pdf', width = 7, height = 7, units = "in")
```


```{r, fig.width = 12, fig.height = 8}
filter(df, tissue %in% common_tissues) %>% ggplot(aes(percent.ribo, ..density..,colour = method)) + 
  geom_freqpoly(binwidth = 0.02) +
  facet_wrap(~ tissue, ncol = 4) + 
  ggtitle("Percent ribosomal protein")

ggsave('figures/percent_ribo_comparison.pdf', width = 7, height = 7, units = "in")
```

```{r, fig.width = 12, fig.height = 8}

filter(df, tissue %in% common_tissues) %>% ggplot(aes(percent.Rn45s, ..density..,colour = method)) + 
  geom_freqpoly(binwidth = 0.02) +
  facet_wrap(~ tissue, ncol = 4) + 
  ggtitle("Percent Rn45s")

ggsave('figures/percent_rn45s_comparison.pdf', width = 7, height = 7, units = "in")
```

The enrichment of Rn45s and ribosomal genes differ strongly between the two methods.

```{r}
tissue_colors <- read_csv('~/maca/metadata/tissue_colors.csv')
tissue_colors <- rename(tissue_colors, tissue = X1)
```

```{r}
medians <- df %>% group_by(method, tissue, run) %>% 
                  summarize(median_Rn45s = median(percent.Rn45s), 
                            median_ribo = median(percent.ribo),
                            n_cells = n()) 
medians <- left_join(medians, tissue_colors, by = 'tissue')

medians %>% filter(tissue %in% common_tissues,) %>% ggplot(aes(x = median_Rn45s, y = median_ribo, color = color, shape = method)) + geom_point()  +
  scale_color_identity(breaks = tissue_colors$color,
                       labels = tissue_colors$tissue,
                       guide = "legend") + 
  ggtitle("Comparing Rn45s and ribo genes between channels and plates.")

ggsave('figures/rn45s_vs_percent_ribo.pdf', width = 7, height = 7, units = "in")

```




##  Dynamic Range of individual cells

We compare individual cells from tissues containing similar numbers of genes per cell.


Get a few cells from each method with values in a certain range.
```{r}
subset.size <- 1
set.seed(1)
cell.subset <- filter(df, tissue == 'Heart', nGene > 2900, nGene < 3100) %>% 
    group_by(method) %>% 
    sample_n(subset.size) %>%
    pull(cell)
```

It's difficult to compare UMI and reads apples to apples. Here both are scaled to 10k reads/UMIs then log(1+x)'d.

```{r, fig.width = 4, fig.height = 8}
pdf('figures/heart_cell_comparison.pdf')
par(mfrow=c(2,1))

expr <- tissX@data[,cell.subset[2]]
expr <- expr[expr > 0]
plot(sort(expr), ylab = "log expression", xlab = "gene index", col = "red")
title(main = "Heart tenx cell")

expr <- tiss@data[,cell.subset[1]]
expr <- expr[expr > 0]
plot(sort(expr), ylab = "log expression", xlab = "gene index", col = "green")
title(main = "Heart plateseq cell")

dev.off()
```

If we instead scaled them to 10^6 both, we would see a different shape for the plate data, with more dynamic range in the low values.

```{r, fig.width = 4, fig.height = 8}
pdf('figures/heart_cell_comparison.pdf')


par(mfrow=c(2,1))

expr <- adjust_lognorm(tissX@data[,cell.subset[2]], 1)
expr <- expr[expr > 0]
plot(sort(expr), ylab = "log expression", xlab = "gene index", col = "red")
title(main = "Heart tenx cell")

expr <- adjust_lognorm(tiss@data[,cell.subset[1]], 100)
expr <- expr[expr > 0]
plot(sort(expr), ylab = "log expression", xlab = "gene index", col = "green")
title(main = "Heart plateseq cell")

dev.off()
```


## Comparing the Average (T cell)

```{r}
anno_plate = read_csv('~/maca/metadata/maca_3month_combined_cell_annotations.csv')
anno_plate <- rename(anno_plate, cell = X1)
#anno_plate <- rename(anno, anno_tissue = tissue)
anno_plate %>% group_by(tissue, annotation, subannotation) %>% summarize(count = n())

anno_tenx = read_csv('~/maca/metadata/maca_3month_combined_cell_annotations_10x.csv')
anno_tenx <- rename(anno_tenx, cell = X1)
#anno_plate <- rename(anno, anno_tissue = tissue)
anno_tenx %>% group_by(tissue, annotation, subannotation) %>% summarize(count = n())
```


```{r}
library(scales)
```

```{r}
plate_v_tenx <- function(tissue_var, annotation_var){

cells_plate <- anno_plate %>% filter(annotation == annotation_var & tissue == tissue_var) %>% pull(cell)
pct_plate <- Matrix::rowMeans(tiss@raw.data[,cells_plate] > 0)
pct_plate
cells_tenx <- anno_tenx %>% filter(annotation == annotation_var & tissue == tissue_var) %>% pull(cell)
pct_tenx <- Matrix::rowMeans(tissX@raw.data[,cells_tenx] > 0)

pct1 = as_data_frame(pct_tenx)
pct1['gene'] <- names(pct_tenx)
pct1['method'] <- 'tenx'
pct1 <- filter(pct1, value > 0.01)

pct2 = as_data_frame(pct_plate)
pct2['gene'] <- names(pct_plate)
pct2['method'] <- 'plate'
pct2 <- filter(pct2, value > 0.01)

pct <- bind_rows(pct1, pct2)

pct  %>% spread(key = method, value = value, fill = 0) %>% 
  ggplot(aes(x = plate, y = tenx)) + geom_hex() + geom_abline(slope=1, intercept=0)  + scale_fill_continuous(trans = log10_trans())
}
```

## Examples

```{r}
plate_v_tenx("Kidney", "tubule_cells")
```

```{r}
plate_v_tenx("Bladder", "luminal_cells")
```

```{r}
plate_v_tenx("Bladder", "basal_cells")
```

```{r}
plate_v_tenx("Liver", "hepatocytes") + ggtitle("Liver hepatocytes")
ggsave('figures/Liver_hepatocytes.pdf', width = 7, height = 7, units = "in")
```

```{r}
plate_v_tenx("Kidney", "fibroblasts") + ggtitle("Kidney fibroblasts")
ggsave('figures/Kidney_fibroblasts.pdf', width = 7, height = 7, units = "in")
```

```{r}
plate_v_tenx("Spleen", "b_cells") + ggtitle("Spleen b_cells")
ggsave('figures/Spleen_b_cells.pdf', width = 7, height = 7, units = "in")
```




### Global stats

Let's just go big!

```{r}
anno_plate %>% select(cell, annotation, subannotation, tissue) %>% rename(anno_tissue = tissue) %>% mutate(method = "plate")
anno_tenx %>% select(cell, annotation, subannotation, tissue) %>% rename(anno_tissue = tissue) %>% mutate(method = "tenx")
```

```{r}
all_pct <- tibble(value = double(), gene = character(), method = character(), tissue = character(), annotation = character())

for(tissue_name in unique(anno_plate$tissue)){
  for(annotation_name in filter(anno_plate, tissue == tissue_name) %>% pull(annotation) %>% unique()){
    cells_plate <- anno_plate %>% filter(annotation == annotation_name & tissue == tissue_name) %>% pull(cell)
    pct_plate <- Matrix::rowMeans(tiss@raw.data[,cells_plate] > 0)
    
    pct = as_data_frame(pct_plate)
    pct['gene'] <- names(pct_plate)
    pct['method'] <- 'plate'
    pct['tissue'] <- tissue_name
    pct['annotation'] <- annotation_name
    pct <- filter(pct, value > 0.01)
    
    all_pct <- bind_rows(all_pct, pct)
  }
}

for(tissue_name in unique(anno_tenx$tissue)){
  for(annotation_name in filter(anno_tenx, tissue == tissue_name) %>% pull(annotation) %>% unique()){
    cells_tenx <- anno_tenx %>% filter(annotation == annotation_name & tissue == tissue_name) %>% pull(cell)
    pct_tenx <- Matrix::rowMeans(tissX@raw.data[,cells_tenx] > 0)
    
    pct = as_data_frame(pct_tenx)
    pct['gene'] <- names(pct_tenx)
    pct['method'] <- 'tenx'
    pct['tissue'] <- tissue_name
    pct['annotation'] <- annotation_name
    pct <- filter(pct, value > 0.01)
    
    all_pct <- bind_rows(all_pct, pct)
    }
}
```

```{r}
all_pct <- all_pct %>% mutate(anno_and_tissue = paste(annotation, tissue, sep='-'))

common_anno_tissue <- all_pct %>% group_by(anno_and_tissue, method) %>% summarize(count = n()) %>%
                        ungroup() %>% group_by(anno_and_tissue) %>% summarize(count = n()) %>%
                        filter(count > 1) %>% pull(anno_and_tissue)

pct_comparison <- all_pct %>% filter(anno_and_tissue %in% common_anno_tissue)  %>% spread(key = method, value = value, fill = 0) %>%
        filter(plate > 0.05 | tenx > 0.05) %>% 
        mutate(plate_larger = (plate>tenx),
                   pct_diff = plate - tenx) %>%
        group_by(tissue, anno_and_tissue) %>% summarize(pct_plate_larger = round(mean(plate_larger),2), net_diff = round(mean(pct_diff),2))

rtissue_colors = tissue_colors %>% rename(tissue = tissue)
pct_comparison <- left_join(pct_comparison, rtissue_colors, by = 'tissue')

pct_comparison %>% ggplot(aes(pct_plate_larger, net_diff)) + geom_vline(xintercept = 0.5) + geom_hline(yintercept = 0) +
   geom_point(aes(color = color), size=2) +
   scale_color_identity(breaks = rtissue_colors$color, 
                        labels = rtissue_colors$tissue, 
                        guide = "legend") + 
  labs(x = 'Percent of genes expressed more in plates', y = 'Mean difference in expression percent') +
  guides(color = guide_legend(override.aes = list(size=2)))

```

```{r}
ggsave('figures/pct_of_pct_plate_v_tenx_by_cell_type.pdf', width = 12, height = 7, units = "in")
```


```{r}
pct_comparison %>% filter(pct_plate_larger < 0.5)
```


## Miscellaneous




```{r}
df %>% group_by(annotation.y, tissue, plate.barcode.x) %>% summarize(count = n()) %>% ungroup() %>% group_by(annotation.y, tissue) %>% summarize(n_plates = n()) %>% arrange(desc(n_plates))

df %>% group_by(annotation.y, tissue) %>% summarize(count = n()) %>% ungroup() %>% group_by(annotation.y) %>% summarize(n_tissues = n()) %>% arrange(desc(n_tissues))
```



```{r}
df %>% group_by(annotation.y, subannotation.y) %>% summarize(count = n()) %>% ungroup()
```


```{r}
df %>% group_by(annotation.y, tissue, plate.barcode.x) %>% summarize(count = n()) %>% ungroup() %>% group_by(annotation.y, tissue) %>% summarize(n_plates = n()) %>% arrange(desc(n_plates))
```

